Figures for Paper

### 2021-09-01 paper_figures.Rmd
Adding in the raw data from different guides - salmon data

rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "UNC13A",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B") +
    xlab("") +
    theme(legend.position = "none") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent) +
    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = source, y= 1.35,label = padj_plot),size = 8) +
    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)

@@

clean_data_table %>% 
    filter(tissue_clean %in% c("Frontal_Cortex", 
                               "Lumbar_Spinal_Cord", 
                               "Cervical_Spinal_Cord", 
                               "Motor_Cortex", 
                               "Temporal_Cortex")) %>% 
    unique() %>% 
    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
    ggplot(aes(x = disease_group2,
               y = IRratioUNC13A)) +
    geom_boxplot(aes(fill = disease_group2)) + 
    facet_wrap(~tissue_clean,scales = 'free_x') +
    stat_pvalue_manual(stat_test_a %>% filter(group1 == "Control" & group2 == "FTLD-TDP"), 
                       label = "p.signif",y.position = 0.98) +
    stat_pvalue_manual(
                       stat_test_a %>% filter(group1 == "ALS-TDP" & group2 == "Control"), 
                       label = "p.signif",y.position = 0.88) +
    ylab("UNC13A IRratio") +
  xlab(element_blank())+
  scale_fill_manual(values = force_colors) +
  ggpubr::theme_pubr() + 
  theme(legend.position = 'none')

@@

```{r unc13b_ir_in_newyork}
clean_data_table %>% 
    filter(tissue_clean %in% c("Frontal_Cortex", 
                               "Lumbar_Spinal_Cord", 
                               "Cervical_Spinal_Cord", 
                               "Motor_Cortex", 
                               "Temporal_Cortex")) %>% 
    unique() %>% 
    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
    ggplot(aes(x = disease_group2,
               y = IRratioUNC13B)) +
    geom_boxplot(aes(fill = disease_group2)) + 
    facet_wrap(~tissue_clean,scales = 'free_x') +
    stat_pvalue_manual(stat_test_b %>% filter(group1 == "Control" & group2 == "FTLD-TDP"), 
                       label = "p.signif",y.position = 0.98) +
    stat_pvalue_manual(
        stat_test_b %>% filter(group1 == "ALS-TDP" & group2 == "Control"), 
        label = "p.signif",y.position = 0.88) +
    ylab("UNC13B IRratio") +
    scale_fill_manual(values = force_colors) +
    ggpubr::theme_pubr() + 
    theme(legend.position = 'none')

@@

@@

Read in pre-processed data (see separate script) and plot
```{r warning=FALSE, message=FALSE}
df <- read_csv(paste0(here(),
                      "/data/nanopore/combined_nanopore_dataframe.csv"))

>>>>

rel_rna_cryptic_amount %>%
  left_join(text_table[,.(plot_name,source)]) %>% 
  unique() %>% 
  mutate(plot_name = fct_relevel(plot_name,"iPSC MN","SH-SY5Y","I3 Neurons","SK-N-DZ_a")) %>% 
    ggbarplot(,
              x = "plot_name",
              add = c("mean_se","jitter"),
              y = "UNC13B",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B") +
    xlab("") +
    theme(legend.position = "none") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent,expand = expansion(mult = c(0, .1))) +
    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = plot_name, y= 1.35,label = padj_plot),size = 8) +
    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = plot_name, y= 1.2,label = log2FoldChange_plot),size = 8)

@@

clean_data_table %>% 
    filter(tissue_clean %in% c("Frontal_Cortex", 
                               "Lumbar_Spinal_Cord", 
                               "Cervical_Spinal_Cord", 
                               "Motor_Cortex", 
                               "Temporal_Cortex")) %>% 
    unique() %>% 
    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
    ggplot(aes(x = disease_group2,
               y = IRratioUNC13A)) +
    geom_boxplot(aes(fill = disease_group2),outlier.shape = NA) + 
   geom_jitter(aes(fill = disease_group2),
               pch = 21,height = 0,
               width = 0.2,size = 3,alpha = 0.8) +
    facet_wrap(~tissue_clean,scales = 'free_x') +
    stat_pvalue_manual(stat_test_a %>% filter(group1 == "Control" & group2 == "FTLD-TDP"), 
                       label = "p.signif",y.position = 0.98) +
    stat_pvalue_manual(
                       stat_test_a %>% filter(group1 == "ALS-TDP" & group2 == "Control"), 
                       label = "p.signif",y.position = 0.88) +
    ylab("UNC13A IRratio") +
  xlab(element_blank())+
  scale_fill_manual(values = force_colors) +
  ggpubr::theme_pubr() + 
  theme(legend.position = 'none')

@@

```{r unc13b_ir_in_newyork}
clean_data_table %>% 
    filter(tissue_clean %in% c("Frontal_Cortex", 
                               "Lumbar_Spinal_Cord", 
                               "Cervical_Spinal_Cord", 
                               "Motor_Cortex", 
                               "Temporal_Cortex")) %>% 
    unique() %>% 
    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
    ggplot(aes(x = disease_group2,
               y = IRratioUNC13B)) +
    geom_boxplot(aes(fill = disease_group2),outlier.shape = NA) + 
   geom_jitter(aes(fill = disease_group2),
               pch = 21,height = 0,
               width = 0.2,size = 3,alpha = 0.8) +    facet_wrap(~tissue_clean,scales = 'free_x') +
    stat_pvalue_manual(stat_test_b %>% filter(group1 == "Control" & group2 == "FTLD-TDP"), 
                       label = "p.signif",y.position = 0.98) +
    stat_pvalue_manual(
        stat_test_b %>% filter(group1 == "ALS-TDP" & group2 == "Control"), 
        label = "p.signif",y.position = 0.88) +
    ylab("UNC13B IRratio") +
  xlab(element_blank()) +
    scale_fill_manual(values = force_colors) +
    ggpubr::theme_pubr() + 
    theme(legend.position = 'none')

@@

# UNC13B isoform expression in-vitro and across the NYGC

```{r unc13b_isoform_cell_lines}
iso_colors = c("#1972A9", "#28A919", "#F49F0B")
names(iso_colors) = c("UNC13B-207","UNC13B-208","UNC13B-210")
#207 -- 1972A9 - normal
#208 -- 28A919 - NMD
#210 -- F49F0B - brain
t_type = fread(file.path(here::here(),"data","transcripts-Summary-Homo_sapiens_Gene_Summary_ENSG00000198722.csv"))

cell_lines = fread(file.path(here::here(),"data","unc13b_salmon_cell_lines.csv"))
cell_lines %>% 
    mutate(sample = gsub(".Aligned.sorted.out","",sample)) %>% 
    left_join(rel_rna_cryptic_amount[,.(condition,plot_name,sample)]) %>% 
    left_join(t_type, by = c("Name" = "Transcript ID")) %>% 
    filter(Name.y %in% c("UNC13B-207",
                         "UNC13B-208",
                         "UNC13B-210")) %>%
    ggplot(aes(x =condition, y = TPM,fill = Name.y)) + 
    geom_boxplot() +
   geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1,jitter.height = 0)) +
    # facet_wrap(~plot_name) +
  ggpubr::theme_pubr() + 
  labs(fill="",x="") + 
  scale_fill_manual(values = iso_colors)

```{r unc13b_isoform_nygc}

unc13b_expression = fread(file.path(here::here(),"data","unc13b_transcript_counts.csv"))

clean_data_table %>% 
    left_join(unc13b_expression) %>% 
    mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
    mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
    select(sample,disease_group2,tissue_clean,unc13a_detected,TPM,Name) %>% 
    filter(!is.na(Name)) %>% 
    unique() %>% 
    filter(tissue_clean %in% c("Frontal_Cortex", 
                               "Lumbar_Spinal_Cord", 
                               "Cervical_Spinal_Cord", 
                               "Motor_Cortex", 
                               "Temporal_Cortex")) %>% 
    left_join(t_type, by = c("Name" = "Transcript ID")) %>% 
        filter(Name.y %in% c("UNC13B-207",
                         "UNC13B-208",
                         "UNC13B-210")) %>%
    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
    ggplot(aes(x = disease_group2, y = TPM, fill = Name.y)) + 
    geom_boxplot() +
    geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1,jitter.height = 0)) +
    facet_wrap(~tissue_clean,scales = 'free') + 
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) + 
    scale_y_continuous(label=comma) +
    ggpubr::theme_pubr() + 
    labs(fill="",x="") +
    scale_fill_manual(values = iso_colors)

@@

Read in pre-processed data (see separate script) and plot
```{r warning=FALSE, message=FALSE}
df <- read_csv(paste0(here::here(),
                      "/data/nanopore/combined_nanopore_dataframe.csv"))
